Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
minor change

aazZ%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Numeric representations and soundness} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and $x_k$. 
The second is the quantization error introduced by the digital controller, 
which performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by the fixed-point representation of numbers~\cite{DBLP:conf/arith/BrainTRW15}. 
% 
Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.

The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed-point domain as above. 
Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T$ where the error $ \delta_T=x\ \%_{c_m}\ \tilde x$, and $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands, namely 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item Division: the operations performed by our controllers in the FWL domain do not include division, so these bounds are inconsequential to this paper. 
\end{enumerate}

@@

In the following, we shall disregard these steady-state errors (caused by FWL effects)
 when the stability is ensured by synthesis, and then verify its safety accounting for the finite-precision errors.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{CEGIS of Safe Controllers for LTI Systems} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

\textit{DC Motor Position and Rate} plants describes the angular position and 
velocity of a DC Motor, respectively. 
\textit{Automotive Cruise System} plant represents the speed of a motor vehicle. 
\textit{Helicopter Longitudinal Motion} plant provides the longitudinal motion model 
of a helicopter. 
\textit{Inverted Pendulum} plant describes a pendulum model
with its center of mass above its pivot point. 
\textit{Magnetic Suspension} plant provides a physical model for which 
a given object is suspended via magnetic field. 
\textit{Magnetized Pointer} plant describes a physical model employed in analogue gauges 
and indicators that is rotated through interaction with magnetic fields.
\textit{1/4 Car Suspension} plant presents a physical model that connects a car to its wheels 
and allows relative motion between both parts.
\textit{Computer Tape Driver} plant describes a physical model to read and write data 
on a storage device in the computer.

Section~\ref{sec:BenchmarksforStateSpace} in the appendix summarizes 
all our benchmarks, which are SISO models (cf. section~\ref{sec:preliminaries}). 
The Inverted Pendulum appears to be a two output system, but it was treated as two 
SISO systems during the experiments. For the sake of simplicity all the states 
measurements are assumed available. In practice, state estimators should be designed 
in order to obtain an estimate of the states from the single output. 
Further extensions of this work can include 
state observers synthesis.

@@

Discretization of an continuous dynamical system turns the ODE into a difference equation, assuming zero-order hold
for the input $\vec{u}$ and continuous integration for the noise $\vec{w}$, to
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d\vec{u}_k + \vec{w}_k,\\
y_k &= \vec{C}_d \vec{x}_ k + \vec{D}_d \vec{u}_ k. 
\end{align}
with covariance for $\vec{w}_k \sim \mathcal{N}(0,Q_d)$,
where
\begin{align}
\label{eq:discretize}
\vec{A}_d &= e^{\vec{A} T_s} = \mathcal{L}^{-1} { ( s \vec{I} - \vec{A} )^{-1} }_{t = T_s},\\
\vec{B}_d &= \int_{0}^{T_s} e^{\vec{A} t} dt\ \vec{B} = \vec{A}^{-1} ( \vec{A}_d - \vec{I} ) \vec{B},\\
\vec{C}_d &= \vec{C},\\
\vec{D}_d &= \vec{D},\\
Q_d &= \int_{0}^{T_s} e^{\vec{A} t} Q e^{\vec{A}^T t} dt.
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT_s)=x_k \wedge y(kT_s) = y_k, \forall k.
\end{align*}
Let $g_d(k)=\mathcal{G}(t,g(t),T)$ be a function that performs the discretization described above, where $g(t)$
represents the continuous dynamics and $g_d(k)$ the corresponding discrete dynamics. 
Let $G(s)=\mathcal{L}(g(t))$ and $G_d(z)=\mathcal{Z}(g_d(k))$ be the corresponding Laplace and Z-transforms
of $g(t)$ and $g_d(k)$. Given this relation, we have 
$$G_d(z)=G(z)|_{z=e^{sT}} : g_d(k)=\mathcal{G}(t,g(t),T) \wedge T < \frac{1}{.5f_s},$$
where $.5f_s$ is the Nyquist frequency of $g(t)$. This last restriction is introduced to avoid the effects of aliasing
which could cause ``phantom poles'' to appear otherwise. 
The eigenvalues of $\vec{A}_d$ corresponding to the poles of $G_d(z)$, and those of $\vec{A}$ corresponding to the poles of $G(s)$ are similarly related.
$\hat{\lambda}_i=e^{-\lambda_iT} : \hat{\lambda}_i \in \sigma(\vec{A}_d) \wedge \lambda_i \in \sigma(\vec{A})$
where $\sigma(\cdot)$ is the spectrum of a matrix.
The following remark is worth mentioning regarding the above.
\begin{remark}
The witnessed maximum amplitude of the discrete signals $x_k,y_k$ may be smaller to that of $x(t),y(t)$ due to synchronism at fraction-frequency sampling.
This means that reasoning about the state-space must consider the effects of this possibility as well of the case of maximal inputs from the continuous specification.
\end{remark}
%  Since \eqref{eq:discretization} is a bi-simulation of \eqref{eq:ode}, we may use the semantics in section \ref{sec:model_semantics} to model continuous dynamical systems.

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Numeric representations and soundness} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and $x_k$. 
The second is the quantization error introduced by the digital controller, 
which performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by the fixed-point representation of numbers. 
% 
Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$. Any mathematical operations performed at the precision $\mathcal{F}_{\langle I,F \rangle}(x)$ will introduce errors, that can be upper bounded~\cite{DBLP:conf/arith/BrainTRW15}. Further details are in Appendix~\ref{appendix:numerical_errors}.

@@

In the following, we shall disregard these steady-state errors (caused by FWL effects)
 when the stability is ensured by synthesis, and then verify its safety accounting for the finite-precision errors. 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{CEGIS of Safe Controllers for LTI Systems} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

\textit{DC Motor Position and Rate} plants describes the angular position and 
velocity of a DC Motor, respectively. 
\textit{Automotive Cruise System} plant represents the speed of a motor vehicle. 
\textit{Helicopter Longitudinal Motion} plant provides the longitudinal motion model 
of a helicopter. 
\textit{Inverted Pendulum} plant describes a pendulum model
with its center of mass above its pivot point. 
\textit{Magnetic Suspension} plant provides a physical model for which 
a given object is suspended via magnetic field. 
\textit{Magnetized Pointer} plant describes a physical model employed in analogue gauges 
and indicators that is rotated through interaction with magnetic fields.
\textit{1/4 Car Suspension} plant presents a physical model that connects a car to its wheels 
and allows relative motion between both parts.
\textit{Computer Tape Driver} plant describes a system to read and write data 
on a storage device in the computer.

Section~\ref{sec:BenchmarksforStateSpace} in the appendix summarizes 
all our benchmarks, which are SISO models (cf. section~\ref{sec:preliminaries}). 
The Inverted Pendulum appears to be a two output system, but it was treated as two 
SISO systems during the experiments. For the sake of simplicity all the states 
measurements are assumed available. In practice, state estimators should be designed 
in order to obtain an estimate of the states from the single output. 
Further work can include 
state observers synthesis.

@@

Discretization of an continuous dynamical system turns the ODE into a difference equation, assuming zero-order hold
for the input $\vec{u}$ and continuous integration for the noise $\vec{w}$, to
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d\vec{u}_k + \vec{w}_k,\\
y_k &= \vec{C}_d \vec{x}_ k + \vec{D}_d \vec{u}_ k. 
\end{align}
with covariance for $\vec{w}_k \sim \mathcal{N}(0,Q_d)$,
where
\begin{align}
\label{eq:discretize}
\vec{A}_d &= e^{\vec{A} T_s} = \mathcal{L}^{-1} { ( s \vec{I} - \vec{A} )^{-1} }_{t = T_s},\\
\vec{B}_d &= \int_{0}^{T_s} e^{\vec{A} t} dt\ \vec{B} = \vec{A}^{-1} ( \vec{A}_d - \vec{I} ) \vec{B},\\
\vec{C}_d &= \vec{C},\\
\vec{D}_d &= \vec{D},\\
Q_d &= \int_{0}^{T_s} e^{\vec{A} t} Q e^{\vec{A}^T t} dt.
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT_s)=x_k \wedge y(kT_s) = y_k, \forall k.
\end{align*}
Let $g_d(k)=\mathcal{G}(t,g(t),T)$ be a function that performs the discretization described above, where $g(t)$
represents the continuous dynamics and $g_d(k)$ the corresponding discrete dynamics. 
Let $G(s)=\mathcal{L}(g(t))$ and $G_d(z)=\mathcal{Z}(g_d(k))$ be the corresponding Laplace and Z-transforms
of $g(t)$ and $g_d(k)$. Given this relation, we have 
$$G_d(z)=G(z)|_{z=e^{sT}} : g_d(k)=\mathcal{G}(t,g(t),T) \wedge T < \frac{1}{.5f_s},$$
where $.5f_s$ is the Nyquist frequency of $g(t)$. This last restriction is introduced to avoid the effects of aliasing
which could cause ``phantom poles'' to appear otherwise. 
The eigenvalues of $\vec{A}_d$ corresponding to the poles of $G_d(z)$, and those of $\vec{A}$ corresponding to the poles of $G(s)$ are similarly related.
$\hat{\lambda}_i=e^{-\lambda_iT} : \hat{\lambda}_i \in \sigma(\vec{A}_d) \wedge \lambda_i \in \sigma(\vec{A})$
where $\sigma(\cdot)$ is the spectrum of a matrix.
The following remark is worth mentioning regarding the above.
\begin{remark}
The witnessed maximum amplitude of the discrete signals $x_k,y_k$ may be smaller to that of $x(t),y(t)$ due to synchronism at fraction-frequency sampling.
This means that reasoning about the state-space must consider the effects of this possibility as well of the case of maximal inputs from the continuous specification.
\end{remark}

\subsection{Errors due to numerical representation} \label{appendix:numerical_errors}

Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.
The following approximation errors will arise in mathematical operations and representation:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed-point domain as above. 
Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T$ where the error $ \delta_T=x\ \%_{c_m}\ \tilde x$, and $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands, namely 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item Division: the operations performed by our controllers in the FWL domain do not include division. However, we do use division in computations at the precision of the plant. $\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3| \leq $, where $c_m=2^{-F}$ as above.
\end{enumerate}

%  Since \eqref{eq:discretization} is a bi-simulation of \eqref{eq:ode}, we may use the semantics in section \ref{sec:model_semantics} to model continuous dynamical systems.
